Structured Versus Decorated Cospans

### 2021-06-20 cospans_new.tex
Completed proof of Theorem 2.5, but see stuff in red.

\begin{thm}
\label{thm:functoriality}
Given categories $\A$ and $\A'$ with finite colimits, lax monoidal pseudofunctors $F, F' \maps (\A',+) \to (\Cat,\times)$, a finite colimit preserving functor $H \maps \A \to \A'$, a lax monoidal pseudofunctor $E \maps (\Cat,\times) \to (\Cat,\times)$ and a {\john monoidal???} natural isomorphism $\theta$ as in the following diagram:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$\A$};
\node (B) at (1,0) {$\Cat$};
\node (C) at (0,-1) {$\A'$};
\node (D) at (1,-1) {$\Cat$};
\node (E) at (0.5,-0.5) {$\Swarrow \theta$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$F$} (B)
(A) edge node[left]{$H$} (C)
(B) edge node[right]{$E$} (D)
(C) edge node[below]{$F'$} (D);
\end{tikzpicture}
\] 
we obtain a double functor $\lH \maps F\lCsp \to F'\lCsp$ defined as above.
\end{thm}

@@

To make $G$ into a strong monoidal functor, we need to equip it with a natural isomorphism
\[  \gamma \maps G(a \xrightarrow{i} b, x) \otimes G(a' \xrightarrow{i'} b', x') \; 
\xrightarrow{\sim} \; G((a \xrightarrow{i} b, x) \otimes (a' \xrightarrow{i'} b', x')).\]
Note that
\[   G(a \xrightarrow{i} b, x) \otimes G(a' \xrightarrow{i'} b', x') = 
\left(H(a) + H(a') \xrightarrow{H(i) + H(i')} H(b) + H(b') , \; \phi'_{H(b), H(b')}(\underline{x},
\underline{x}')\right) \]
where $\phi'$ is the laxator for $F'$, while
\[    G((a \xrightarrow{i} b, x) \otimes (a' \xrightarrow{i'} b', x')) =
\left(H(a+a') \xrightarrow{H(i+i')} H(b+b'), \; \underline{\phi_{b,b'}(x,x')} \right)
\]
where $\phi$ is the laxator for $F$.  Since $H$ preserves finite colimits, the isomorphism
$\gamma$ is obvious except

The laxator is given by the following natural isomorphism between $\theta_m E(p) \phi \times \theta_n E(q) \phi$ and $\theta_{m+n} E(p \times q) \phi$:
\[
\begin{tikzpicture}[scale=1.5]
\hspace{-.6in}
\node () at (3,-0.5) {$\scriptscriptstyle{\stackrel{E_{p,q}} \cong}$};
\node () at (1.25,-0.5) {$\scriptscriptstyle{\stackrel{\lambda} \cong}$};
\node (L1) at (0.5,0) {$\scriptscriptstyle{1 \times E(1)}$};
\node (L2) at (0.5,-1) {$\scriptscriptstyle{E(1)}$};
\node (K) at (-1,0) {$\scriptscriptstyle{1 \times 1}$};
\node (K2) at (-1,-1) {$\scriptscriptstyle{1}$};
\node (B) at (2,0) {$\scriptscriptstyle{E(1) \times E(1)}$};
\node (D) at (2,-1) {$\scriptscriptstyle{E(1 \times 1)}$};
\node (E) at (4,0) {$\scriptscriptstyle{E(F(m)) \times E(F(n))}$};
\node (F) at (4,-1) {$\scriptscriptstyle{E(F(m) \times F(n))}$};
\node (G) at (7,0) {$\scriptscriptstyle{F'(H(m)) \times F'(H(n))}$};
\node (H) at (10,0) {$\scriptscriptstyle{F'(H(m)+H(n))}$};
\node (I) at (7,-1) {$\scriptscriptstyle{E(F(m+n))}$};
\node (J) at (10,-1) {$\scriptscriptstyle{F'(H(m+n))}$};
\path[->,font=\scriptsize,>=angle 90]
(K) edge node [left] {$\lambda$} (K2)
(K) edge node [above] {$1 \times E_0$} (L1)
(K2) edge node [above] {$E_0$} (L2)
(H) edge node [right] {$F'(H_{m,n})$} (J)
(E) edge node [above] {$\theta_m \times \theta_n$} (G)
(G) edge node [above] {$\phi'_{H(m),H(n)}$} (H) 
(F) edge node [above] {$E(\phi_{m,n})$} (I)
(I) edge node [above] {$\theta_{m+n}$} (J)
(B) edge node [above] {$E(p) \times E(q)$} (E)
(D) edge node [above] {$E(p \times q)$} (F)
(E) edge node [right] {$E_{F(m),F(n)}$} (F)
(L1) edge node[above]{$E_0 \times 1$} (B)
(L1) edge node [left] {$\lambda$} (L2)
(L2) edge node[above]{$E(\lambda^{-1})$} (D)
(B) edge node[right]{$E_{1,1}$} (D);
\end{tikzpicture}
\]
and we also have a natural isomorphism between $0_{\inta F'}$ and $G(0_{\inta F})$ given by:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (-1,0.5) {$\scriptscriptstyle{1}$};
\node () at (4,0.66) {$\scriptscriptstyle{\cong}$};
\node (A1) at (0,0) {$\scriptscriptstyle{E(1)}$};
\node (A2) at (2,0) {$\scriptscriptstyle{E(F(0_\A))}$};
\node (A3) at (4,0) {$\scriptscriptstyle{E(F(0_\A))}$};
\node (A4) at (6,0) {$\scriptscriptstyle{F'(H(0_\A))}$};
\node (A') at (2,0.5) {$\scriptscriptstyle{F'(H(0_\A))}$};
\node (B1) at (0,1) {$\scriptscriptstyle{F'(0_{\A'})}$};
\node (B2) at (6,1) {$\scriptscriptstyle{F'(0_{\A'})}$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node [below] {$\scriptscriptstyle{E_0}$} (A1)
(A) edge node [above] {$\scriptscriptstyle{\phi_0'}$} (B1)
(A1) edge node [above] {$\scriptscriptstyle{E(\phi_0)}$} (A2)
(A2) edge node [below] {$\scriptscriptstyle{E(F(!_{0_\A}))}$} (A3)
(A3) edge node [below] {$\scriptscriptstyle{\theta_{0_\A}}$} (A4)
(A2) edge node [left] {$\scriptscriptstyle{\theta_{0_\A}}$} (A')
(B1) edge node [fill=white] {$\scriptscriptstyle{F'(H_0)}$} (A')
(A') edge node [fill=white] {$\scriptscriptstyle{F'(H(!_{0_\A}))}$} (A4)
(B2) edge node [right] {$\scriptscriptstyle{F'(H_0)}$} (A4)
(B1) edge node [above] {$\scriptscriptstyle{F'(!_{0_{\A'}})}$} (B2);
\end{tikzpicture}
\]
One can show that this laxator and unit morphism satisfy the four required coherence laws of a symmetric strong monoidal functor.
\end{proof}

>>>>

\begin{thm}
\label{thm:functoriality}
Given categories $\A$ and $\A'$ with finite colimits, lax monoidal pseudofunctors $F, F' \maps (\A',+) \to (\Cat,\times)$, a finite colimit preserving functor $H \maps \A \to \A'$, a lax monoidal pseudofunctor $E \maps (\Cat,\times) \to (\Cat,\times)$ and a {\john monoidal???} natural isomorphism $\theta$ as in the following diagram:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$\A$};
\node (B) at (1,0) {$\Cat$};
\node (C) at (0,-1) {$\A'$};
\node (D) at (1,-1) {$\Cat$};
\node (E) at (0.5,-0.5) {$\Swarrow \theta$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$F$} (B)
(A) edge node[left]{$H$} (C)
(B) edge node[right]{$E$} (D)
(C) edge node[below]{$F'$} (D);
\end{tikzpicture}
\] 
we obtain a double functor $\lH \maps F\lCsp \to F'\lCsp$ defined as above.  If $F, F'$ and $E$ are symmetric, then $\|H \maps F\lCsp \to F'\lCsp$ is a symmetric monoidal double functor. 
\end{thm}

@@

To make $G$ into a monoidal functor, we need to equip it with a laxator
\[  \gamma \maps G(a \xrightarrow{i} b, x) \otimes G(a' \xrightarrow{i'} b', x') \; 
\to \; G((a \xrightarrow{i} b, x) \otimes (a' \xrightarrow{i'} b', x')).\]
Note that
\[   G(a \xrightarrow{i} b, x) \otimes G(a' \xrightarrow{i'} b', x') = 
\left(H(a) + H(a') \xrightarrow{H(i) + H(i')} H(b) + H(b') , \; \phi'_{H(b), H(b')}(\underline{x},
\underline{x}')\right) \]
where $\phi'$ is the laxator for $F'$, while
\[    G((a \xrightarrow{i} b, x) \otimes (a' \xrightarrow{i'} b', x')) =
\left(H(a+a') \xrightarrow{H(i+i')} H(b+b'), \; \underline{\phi_{b,b'}(x,x')} \right)
\]
where $\phi$ is the laxator for $F$.  Since $H$ preserves finite colimits, the laxator
$\gamma$ is obvious except for the morphism from $\phi'_{H(b), H(b')}(\underline{x},
\underline{x}')$ to $\underline{\phi_{b,b'}(x,x')}$.   This is given by the following natural
transformation:
\[
\begin{tikzpicture}[scale=1.5]
\node () at (3,-0.5) {$\scriptscriptstyle{\stackrel{E_{x,x'}} \cong}$};
\node () at (1.5,-0.5) {$\scriptscriptstyle{\stackrel{\lambda} \cong}$};
\node (L1) at (1,0) {$\scriptscriptstyle{1 \times E(1)}$};
\node (L2) at (1,-1) {$\scriptscriptstyle{E(1)}$};
\node (K) at (0,0) {$\scriptscriptstyle{1 \times 1}$};
\node (K2) at (0,-1) {$\scriptscriptstyle{1}$};
\node (B) at (2,0) {$\scriptscriptstyle{E(1) \times E(1)}$};
\node (D) at (2,-1) {$\scriptscriptstyle{E(1 \times 1)}$};
\node (E) at (4,0) {$\scriptscriptstyle{E(F(b)) \times E(F(b'))}$};
\node (F) at (4,-1) {$\scriptscriptstyle{E(F(b) \times F(b'))}$};
\node (G) at (6,0) {$\scriptscriptstyle{F'(H(b)) \times F'(H(b'))}$};
\node (H) at (8,0) {$\scriptscriptstyle{F'(H(b)+H(b'))}$};
\node (I) at (6,-1) {$\scriptscriptstyle{E(F(b+b'))}$};
\node (J) at (8,-1) {$\scriptscriptstyle{F'(H(b+b'))}$};
\path[->,font=\scriptsize,>=angle 90]
(K) edge node [left] {$\lambda$} (K2)
(K) edge node [above] {$1 \times E_0$} (L1)
(K2) edge node [below] {$E_0$} (L2)
(H) edge node [right] {$F'(H_{b,b'})$.} (J)
(E) edge node [above] {$\theta_b \times \theta_{b'}$} (G)
(G) edge node [above] {$\phi'_{H(b),H(b')}$} (H) 
(F) edge node [below] {$E(\phi_{b,b'})$} (I)
(I) edge node [below] {$\theta_{b+b'}$} (J)
(B) edge node [above] {$E(x) \times E(x')$} (E)
(D) edge node [below] {$E(x \times x')$} (F)
(E) edge node [right] {$E_{F(b),F(b')}$} (F)
(L1) edge node[above]{$E_0 \times 1$} (B)
(L1) edge node [left] {$\lambda$} (L2)
(L2) edge node[below]{$E(\lambda^{-1})$} (D)
(B) edge node[right]{$E_{1,1}$} (D);
\end{tikzpicture}
\]
We also need a natural transformation expressing lax preservation of the unit object by $G$, which
is defined similarly.   One can show that this laxator and unit morphism satisfy the coherence laws of a lax monoidal functor, and they are invertible, so $G$ is strong monoidal functor.   \john{If furthermore $F, F'$ and $E$ are symmetric, one can check that $G$ is symmetric monoidal, and that $\lH$ is a symmetric monoidal double functor.}
\end{proof}

\iffalse
we have a natural isomorphism between $0_{\inta F'}$ and $G(0_{\inta F})$ given by:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (-1,0.5) {$\scriptscriptstyle{1}$};
\node () at (4,0.66) {$\scriptscriptstyle{\cong}$};
\node (A1) at (0,0) {$\scriptscriptstyle{E(1)}$};
\node (A2) at (2,0) {$\scriptscriptstyle{E(F(0_\A))}$};
\node (A3) at (4,0) {$\scriptscriptstyle{E(F(0_\A))}$};
\node (A4) at (6,0) {$\scriptscriptstyle{F'(H(0_\A))}$};
\node (A') at (2,0.5) {$\scriptscriptstyle{F'(H(0_\A))}$};
\node (B1) at (0,1) {$\scriptscriptstyle{F'(0_{\A'})}$};
\node (B2) at (6,1) {$\scriptscriptstyle{F'(0_{\A'})}$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node [below] {$\scriptscriptstyle{E_0}$} (A1)
(A) edge node [above] {$\scriptscriptstyle{\phi_0'}$} (B1)
(A1) edge node [above] {$\scriptscriptstyle{E(\phi_0)}$} (A2)
(A2) edge node [below] {$\scriptscriptstyle{E(F(!_{0_\A}))}$} (A3)
(A3) edge node [below] {$\scriptscriptstyle{\theta_{0_\A}}$} (A4)
(A2) edge node [left] {$\scriptscriptstyle{\theta_{0_\A}}$} (A')
(B1) edge node [fill=white] {$\scriptscriptstyle{F'(H_0)}$} (A')
(A') edge node [fill=white] {$\scriptscriptstyle{F'(H(!_{0_\A}))}$} (A4)
(B2) edge node [right] {$\scriptscriptstyle{F'(H_0)}$} (A4)
(B1) edge node [above] {$\scriptscriptstyle{F'(!_{0_{\A'}})}$} (B2);
\end{tikzpicture}
\]
\fi
